---
title: "Categorical Data Analysis Exercise Solutions"
output:
  pdf_document:
  html_document:
    df_print: paged
---

# Set-up
```{r setup, message=FALSE, warning=FALSE}
library(tidyverse) # for tidyverse
library(here) # for file paths
library(survey) # for survey analysis
library(srvyr) # for tidy survey analysis

anes <- read_rds(here("Data", "anes_2020.rds")) %>%
   mutate(Weight=Weight/sum(Weight)*231592693) 
# adjust weight to sum to citizen pop, 18+ in Nov 2020 per ANES methodology documentation
anes_des <- anes %>%
   as_survey_design(weights = Weight,
                    strata = Stratum,
                    ids = VarUnit,
                    nest = TRUE)
```

1. How many females have a graduate degree? Hint: the variables `Gender` and `Education` will be useful.

```{r ex1_1}
#Option 1:
femgd <- anes_des %>%
   filter(Gender=="Female", Education=="Graduate") %>%
   survey_count(name="n")
#Option 2:
femgd <- anes_des %>%
   filter(Gender=="Female", Education=="Graduate") %>%
   summarize(
      N=survey_total(),  .groups="drop"
   )

```

There are `r formatC(pull(femgd, N), format="d", big.mark=",")` females with a graduate degree.


2. What percentage of people identify as "Strong democrat"? Hint: The variable `PartyID` indicates what party people identify with.

```{r ex1_2}
(psd <- anes_des %>%
   group_by(PartyID) %>% 
   summarize(
      p=survey_mean()
   ) %>%
   filter(PartyID=="Strong democrat"))
```

`r str_c(round(pull(psd, p)*100, 1), "%")` of people identify as a strong democrat.


3. What percentage of people who voted in the 2020 election identify as "Strong republican"? Hint: The variable `VotedPres2020` indicates whether someone voted in 2020.

```{r ex1_3}
(psr <- anes_des %>%
   filter(VotedPres2020=="Yes") %>% 
   group_by(PartyID) %>% 
   summarize(
      p=survey_mean()
   ) %>%
   filter(PartyID=="Strong republican"))
```

`r str_c(round(pull(psr, p)*100, 1), "%")` of people identify as a strong republican among those who voted in 2020.

4. What percentage of people voted in both the 2016 election and in the 2020 election?  Include the logit confidence interval. Hint: The variable `VotedPres2016` indicates whether someone voted in 2016.

```{r ex1_4}
(pvb <- anes_des %>%
   filter(!is.na(VotedPres2016), !is.na(VotedPres2020)) %>%
   group_by(interact(VotedPres2016, VotedPres2020)) %>% 
   summarize(
      p=survey_prop(var="ci", method="logit"),
   ) %>%
   filter(VotedPres2016=="Yes", VotedPres2020=="Yes"))
```

`r str_c(round(pull(pvb, p)*100, 1), "%")` (`r round(pull(pvb, p_low)*100, 1)`-`r str_c(round(pull(pvb, p_upp)*100, 1), "%")`) voted in both the 2016 and 2020 elections. 



5. What is the design effect for the proportion of people who voted early? Hint: The variable `EarlyVote2020` indicates whether someone voted early in 2020.

```{r ex1_5}
(pdeff <- anes_des %>%
   filter(!is.na(EarlyVote2020)) %>%
   group_by(EarlyVote2020) %>%
   summarize(
      p=survey_mean(deff=TRUE)
   ) %>%
   filter(EarlyVote2020=="Yes"))
```

The design effect is `r round(pull(pdeff, p_deff), 2)`.

6. Were people who lean democrat more likely to vote early in the 2020 election? Use a logistic regression.

```{r ex1_6}
anes_des %>%
   filter(!is.na(PartyID), !is.na(EarlyVote2020)) %>%
   group_by(PartyID, EarlyVote2020) %>%
   summarise(
      p=survey_mean(),
      .groups="drop"
   ) %>%
   filter(EarlyVote2020=="Yes")

(pid_vote <- anes_des %>%
   svyglm(design=.,
          formula=(EarlyVote2020=="Yes")~PartyID,
          family=quasibinomial(),
          na.action=na.omit))
summary(pid_vote)
```

Strong Democrats are more likely to vote early.

# Bonus

1. What percentage of people lean republican? These are individuals that are strong republicans, not very strong republicans and are independent-republicans. Include an appropriate confidence interval. Hint: to get the correct confidence interval, create a new variable BEFORE calculating the estimate.

```{r exb_1}

#Solution 1: Using forcats package
anes_des %>%
   filter(!is.na(PartyID)) %>%
   mutate(PartyID3=fct_collapse(PartyID,
                                LeanDem=c("Strong democrat",
                                          "Not very strong democrat",
                                          "Independent-democrat"),
                                LeanRep=c("Strong republican",
                                          "Not very strong republican",
                                          "Independent-republican"),
                                other_level="Other")) %>% 
   group_by(PartyID3) %>% 
   summarize(p=survey_prop(vartype="ci", proportion = TRUE))

#Solution 2: Using case_when
anes_des %>%
   filter(!is.na(PartyID)) %>%
   mutate(PartyID3=case_when(PartyID %in% c("Strong democrat",
                                            "Not very strong democrat",
                                            "Independent-democrat")~"LeanDem",
                             PartyID %in% c("Strong republican",
                                            "Not very strong republican",
                                            "Independent-republican")~"LeanRep",
                             TRUE~"Other")) %>% 
   group_by(PartyID3) %>% 
   summarize(p=survey_prop(vartype="ci", proportion = TRUE))

```

# Session information

```{r si}
devtools::session_info(pkgs="attached")
```

